---
title: "Bus Ridership Analysis"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

#### Create a vector s1 whose elements correspond to the number of passengers who can possibly get on board the bus at a particular stop.P[1,P]

```{r}
s1 = c(0,1,2)
```

#### Create a vector p1 whose elements correspond to the respective chances of the number of passengers identified in the previous question getting on board the bus at a particular stop.

```{r}
p1 = c(0.5,0.4,0.1)
```

#### Create a matrix 2x10 matrix P such that the jth column of this matrix should have a passenger on the bus’s continuing-on and getting-off chances for the jth stop for j = 1, . . . , 10.

```{r}
P = matrix(c(0.99,0.01, 0.2,0.8, 0.2,0.8, 0.8,0.2, 0.8,0.2, 0.8,0.2, 0.8,0.2, 0.2,0.8, 0.2,0.8, 0,1), nrow=2, ncol=10)
```

#### Create a vector s2 whose elements correspond to the encoded versions of a passenger staying on the bus and getting off of the bus at a particular stop, respectively

```{r}
s2 = c(0,1)
```

```{r}
sample(s1, size=1, replace = F, prob=p1)
```

```{r}
n = 30 # zero passengers in the bus at stop-1
for (j in c(1:9)){
  # simulating the continuing on or getting off of the bus
  if (n > 0){ # if there is at least one passenger in the bus
    # Updating the current passengers in the bus by subtracting off the number of passengers that get off the bus at the j-th stop
    n = n - sum(sample(s2, size=n, replace=T, prob=P[,j]))
  }
  # update the number of passengers in the bus by adding the number of passenger into the bus at the j-th stop
n = n + sample(s1, size=1, replace=F, prob=p1)
cat(sprintf('No, of passenger after stop %d = %d\n', j,n))
}





```



